---
title: "Time Series Analysis Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=TRUE,results='hide',fig.keep='all',message=FALSE}
library(fpp)
library(tseries)
library(TSA)
library(ggplot2)
library(forecast)
library(expsmooth)
library(readxl)
library(xts)
library(plotrix)
```

Reading in data:
```{r}
datapath <- "/Users/chrisolen/Documents/uchicago_courses/time_series/project"
solar <- ts(read.csv(paste(datapath, "solar.csv",sep = '/')), start=c(1989,1), end=c(2019,4), frequency=12)
solar_production <- solar[,2]
solar_production_train <- ts(solar_production[1:348],start=c(1989,1), frequency=12)
solar_production_test <- ts(solar_production[349:360], end=c(2018,12), frequency=12)
head(solar)
```

Plotting the series:
```{r}
plot(solar_production)
plot(decompose(solar_production, type = c("multiplicative")))
```

BoxCox.lambda recommends a lambda of -0.02 (roughly a log transformation):
```{r}
lambda <- BoxCox.lambda(solar_production_train)
lambda
plot(BoxCox(solar_production, lambda = lambda))
```

# Baseline Model:

$Y_t=Y_{t-1}+e_t$

```{r}
baseline <- Arima(solar_production_train, order = c(0,1,0), lambda = "auto")
summary(baseline)
checkresiduals(baseline)
```

# Performing CV and calculating MAE, MSE, and AICc for baseline model:

```{r}
k <- 180 # Minimum data length for fitting a model
n <- length(solar_production_train) # Number of data points
p <- 12 # Period
H <- 12 # Forecast Horizon

st <- tsp(solar_production_train)[1]+(k-2)/p # start time in units
print(st)
model_error <- matrix(NA,n-k,H)
AICc_model <- list()

### One Month rolling forecasting ###
  
# Sliding Window - With window of length = 180

for(i in 1:(n-k))
{

  train_sli <- window(solar_production_train, start=st+(i-k+1)/p, end=st+i/p) ## Window Length: k
  val <- window(solar_production_train, start=st + (i+1)/p, end=st + (i+H)/p) ## Window Length: H
  if (i<4) {
  cat(c("*** CV", i,":","len(Sliding Window):",length(train_sli), "len(Val):",length(val),'\n'  ))
  cat(c("*** TRAIN - Sliding WIndow:",tsp(train_sli)[1],'-',tsp(train_sli)[2],'\n'))
  cat(c("*** TEST:",tsp(val)[1],'-',tsp(val)[2],'\n'))
  cat("*************************** \n \n")
  }
  
  fit_model <- Arima(train_sli, order=c(0,1,0), seasonal=list(order=c(0,0,0), period=p),
                include.drift=TRUE, lambda='auto', method="ML")
  
  fcast_model <- forecast(fit_model, h=H)
  
  model_error[i,1:length(val)] <- fcast_model[['mean']]-val
  
  AICc_model <- append(AICc_model, fit_model$aicc)
}
```
```{r}
MAE <- function(x){
  return(mean(abs(x),na.rm=TRUE))
}

MSE <- function(x){
  return(mean(x^2, na.rm=TRUE))
}

flattenlist <- function(x){  
  morelists <- sapply(x, function(xprime) class(xprime)[1]=="list")
  out <- c(x[!morelists], unlist(x[morelists], recursive=FALSE))
  if(sum(morelists)){ 
    Recall(out)
  }else{
    return(out)
  }
}

MAE_model <- apply(model_error, 1, MAE)
MSE_model <- apply(model_error, 1, MSE)

baseline.MAE <- mean(MAE_model)
baseline.MSE <- mean(MSE_model)
results <- rbind(baseline.MAE, baseline.MSE)
rownames(results) <- c("MAE","MSE")
colnames(results) <- c("Baseline")
cat("Mean AICc:", mean(unlist(flattenlist(AICc_model))))

```
```{r}
results

```

# Forecasting Error for Baseline Model:

```{r}
baseline.forecast.MSE <- mean((forecast(baseline, h = 12)$mean - solar_production_test)^2)
baseline.forecast.MSE
baseline.forecast.MAE <- mean(abs(forecast(baseline, h = 12)$mean - solar_production_test))
baseline.forecast.MAE
plot(forecast(baseline, h = 12))

```

# ARIMA Models:

Plotting ACF and PACF:
```{r}
solar_production_train.transformed <- BoxCox(solar_production_train, lambda = lambda) 
acf1 <- acf(as.vector(solar_production_train.transformed), lag = 24)
plot(acf1, main = "ACF - No Differencing")
pacf1 <- pacf(as.vector(solar_production_train.transformed), lag = 24)
plot(pacf1, main = "PACF - No Differencing")
```


Significant autocorrelation at lags 1 and 4. Seasonal pattern seen in the decay. ACF function indicates non-stationarity. 

Employ Augmented Dickey-Fuller to confirm:
```{r}
adf.test(solar_production_train.transformed, alternative='stationary')
```

Definitely not stationary!

Plotting the unit circle for the baseline model:
```{r}
nodifferencing <- auto.arima(solar_production_train.transformed)
p1 <- c(1,-nodifferencing$ar) # coefficients of the characteristic polynomial
r1 <- polyroot(p1) # and solving for the roots
r1Re <- Re(r1) # finding real parts of the roots
r1Im <- Im(r1) # finding imaginary parts of the roots
plot(r1Re,r1Im,asp=1,xlim=c(min(r1Re),max(r1Re)),ylim=c(min(r1Im),max(r1Im)))
draw.circle(0,0,radius=1)
abline(v=0)
abline(h=0)

```

One of the roots is way outside of the unit circle. 

Let's apply first order differencing and re-examine the ACF and PACF:
```{r}
acf2 <- acf(as.vector(diff(solar_production_train.transformed)), lag = 24)

pacf2 <- pacf(as.vector(diff(solar_production_train.transformed)), lag = 24)
```

There's still a pretty clear seasonal pattern, particularly in the ACF. 

Let's apply seasonal differencing and re-examine the ACF and PACF:
```{r}
acf3 <- acf(as.vector(diff(diff(solar_production_train.transformed), lag = 12)), lag = 24)
pacf3 <- pacf(as.vector(diff(diff(solar_production_train.transformed), lag = 12)), lag = 24)
```


ACF and PACF yield significant values at lag = 1 and after one seasonal period. Let's apply an ARIMA(1,1,1)(1,1,1)12:
```{r}
arima.model.1 <- Arima(solar_production_train, order=c(1,1,1), seasonal=list(order=c(1,1,1), period=12), lambda='auto')
summary(arima.model.1)
checkresiduals(arima.model.1)
```

Let's apply an ARIMA(1,1,1)(0,1,1)12:
```{r}
arima.model.2 <- Arima(solar_production_train, order=c(1,1,1), seasonal=list(order=c(0,1,1), period=12), lambda='auto')
summary(arima.model.2)
checkresiduals(arima.model.2)
```

All point estimates are now significant, but we still reject the null hypothesis that there is no serial autocorrelation in the residuals, per Box-Ljung. 

Let's use EACF to see if we can get a more granular look at AR/MA orders:
```{r}
eacf.model <- eacf(diff(diff(log(solar_production_train), lag = 12)), 12, 12)

compare.with <- 2/sqrt(length(diff(diff(log(solar_production_train), lag = 12))))
print(abs(eacf.model$eacf)-compare.with,digits=2)

```

ARIMA(0,1,1)(0,1,1)12 seems like another candidate:


```{r}
arima.model.3 <- Arima(solar_production_train, order=c(0,1,1), seasonal=list(order=c(0,1,1), period=12), lambda='auto', method="ML")
summary(arima.model.3)
checkresiduals(arima.model.3)
```

All point estimates are still significant and now we do not reject the null hypothesis of the Box-Ljung test at a confidence level of .01.  

One last attempt per EACF: ARIMA(2,1,1)(0,1,1)12:
```{r}
arima.model.4 <- Arima(solar_production, order=c(2,1,1), seasonal=list(order=c(0,1,1), period=12), lambda='auto', method="ML")
summary(arima.model.4)
checkresiduals(arima.model.4)
```

The second order autoregression point estimate is insignificant, the Box-Ljung p-value has dropped below .01, and AICc has increased. 

One last ARIMA exploration before we check Auto.Arima: Let's hone in on the fact that PACF was significant at lag 4 by nullifying all of the order coefficients prior:
```{r}
arima.model.5 <- Arima(solar_production_train, order=c(4,1,1), seasonal=list(order=c(0,1,1), period=12), method="ML", fixed = c(NA,0,0,NA,NA,NA), lambda = 'auto')
arima.model.5
checkresiduals(arima.model.5)
```

The point estimate for order 4 moving average is insignificant, and we don't see much improvement with respect to the Box-Ljung test or AICc. 

Let's employ Auto.Arima:
```{r}
auto.arima.model <- auto.arima(solar_production_train, lambda = 'auto')
summary(auto.arima.model)
checkresiduals(auto.arima.model)
```

We get an ARIMA(0,1,2)(0,1,1)12 from Auto.Arima. 

Let's apply cross-validation on ARIMA(0,1,2)(0,1,1)12 and ARIMA(0,1,1)(0,1,1)12.

# Performing CV and calculating MAE, MSE, and AICc for ARIMA(0,1,1)(0,1,1)12:

$y_t=(1+-0.2384e_{t-1})(1+-0.6214e_{t-12})$

```{r}
k <- 180 # Minimum data length for fitting a model
n <- length(solar_production_train) # Number of data points
p <- 12 # Period
H <- 12 # Forecast Horizon

st <- tsp(solar_production_train)[1]+(k-2)/p # start time in units
print(st)
model_error <- matrix(NA,n-k,H)
AICc_model <- list()

### One Month rolling forecasting ###
  
# Sliding Window - With window of length = 180

for(i in 1:(n-k))
{

  train_sli <- window(solar_production_train, start=st+(i-k+1)/p, end=st+i/p) ## Window Length: k
  val <- window(solar_production_train, start=st + (i+1)/p, end=st + (i+H)/p) ## Window Length: H
  if (i<4) {
  cat(c("*** CV", i,":","len(Sliding Window):",length(train_sli), "len(Val):",length(val),'\n'  ))
  cat(c("*** TRAIN - Sliding WIndow:",tsp(train_sli)[1],'-',tsp(train_sli)[2],'\n'))
  cat(c("*** TEST:",tsp(val)[1],'-',tsp(val)[2],'\n'))
  cat("*************************** \n \n")
  }
  
  fit_model <- Arima(train_sli, order=c(0,1,1), seasonal=list(order=c(0,1,1), period=12), lambda='auto', method="ML")
  
  fcast_model <- forecast(fit_model, h=H)
  
  model_error[i,1:length(val)] <- fcast_model[['mean']]-val
  
  AICc_model <- append(AICc_model, fit_model$aicc)
}
```

```{r}
MAE <- function(x){
  return(mean(abs(x),na.rm=TRUE))
}

MSE <- function(x){
  return(mean(x^2, na.rm=TRUE))
}

flattenlist <- function(x){  
  morelists <- sapply(x, function(xprime) class(xprime)[1]=="list")
  out <- c(x[!morelists], unlist(x[morelists], recursive=FALSE))
  if(sum(morelists)){ 
    Recall(out)
  }else{
    return(out)
  }
}

MAE_model <- apply(model_error, 1, MAE)
MSE_model <- apply(model_error, 1, MSE)

arima011.MAE <- mean(MAE_model)
arima011.MSE <- mean(MSE_model)
results <- cbind(results,rbind(arima011.MAE, arima011.MSE))
colnames(results) <- c("Baseline","ARIMA(0,1,1)(0,1,1)12")

cat("Mean AICc:", mean(unlist(flattenlist(AICc_model))))
```
```{r}
results

```

# Performing CV and calculating MAE, MSE, and AICc for ARIMA(0,1,2)(0,1,1)12:

$y_t=(1+-0.2262e_{t-1}+-0.1053e_{t-2})(1+-0.6104e_{t-12})$

```{r}
k <- 180 # Minimum data length for fitting a model
n <- length(solar_production_train) # Number of data points
p <- 12 # Period
H <- 12 # Forecast Horizon

st <- tsp(solar_production_train)[1]+(k-2)/p # start time in units
print(st)
model_error <- matrix(NA,n-k,H)
AICc_model <- list()

### One Month rolling forecasting ###
  
# Sliding Window - With window of length = 180

for(i in 1:(n-k))
{

  train_sli <- window(solar_production_train, start=st+(i-k+1)/p, end=st+i/p) ## Window Length: k
  val <- window(solar_production_train, start=st + (i+1)/p, end=st + (i+H)/p) ## Window Length: H
  if (i<4) {
  cat(c("*** CV", i,":","len(Sliding Window):",length(train_sli), "len(Val):",length(val),'\n'  ))
  cat(c("*** TRAIN - Sliding WIndow:",tsp(train_sli)[1],'-',tsp(train_sli)[2],'\n'))
  cat(c("*** TEST:",tsp(val)[1],'-',tsp(val)[2],'\n'))
  cat("*************************** \n \n")
  }
  
  fit_model <- Arima(train_sli, order=c(0,1,2), seasonal=list(order=c(0,1,1), period=p), method="ML", lambda = 'auto')
  
  fcast_model <- forecast(fit_model, h=H)
  
  model_error[i,1:length(val)] <- fcast_model[['mean']]-val
  
  AICc_model <- append(AICc_model, fit_model$aicc)
}
```
```{r}
MAE <- function(x){
  return(mean(abs(x),na.rm=TRUE))
}

MSE <- function(x){
  return(mean(x^2, na.rm=TRUE))
}

flattenlist <- function(x){  
  morelists <- sapply(x, function(xprime) class(xprime)[1]=="list")
  out <- c(x[!morelists], unlist(x[morelists], recursive=FALSE))
  if(sum(morelists)){ 
    Recall(out)
  }else{
    return(out)
  }
}

MAE_model <- apply(model_error, 1, MAE)
MSE_model <- apply(model_error, 1, MSE)

arima012.MAE <- mean(MAE_model)
arima012.MSE <- mean(MSE_model)
results <- cbind(results,rbind(arima012.MAE, arima012.MSE))
colnames(results) <- c("Baseline","ARIMA(0,1,1)(0,1,1)12", "ARIMA(0,1,2)(0,1,1)12")

cat("Mean AICc:", mean(unlist(flattenlist(AICc_model))))
```
```{r}
results

```

# Forecasting Error for Arima Model:

```{r}
arima011.forecast.MSE <- mean((forecast(arima.model.3, h = 12)$mean - solar_production_test)^2)
arima011.forecast.MSE
arima011.forecast.MAE <- mean(abs(forecast(arima.model.3, h = 12)$mean - solar_production_test))
arima011.forecast.MAE
plot(forecast(arima.model.3, h = 12))
```
\
# Forecasting Error for Auto Arima:

```{r}
arima012.forecast.MSE <- mean((forecast(auto.arima.model, h = 12)$mean - solar_production_test)^2)
arima012.forecast.MSE
arima012.forecast.MAE <- mean(abs(forecast(auto.arima.model, h = 12)$mean - solar_production_test))
arima012.forecast.MAE
plot(forecast(auto.arima.model, h = 12))

```

# ETS:
$y_t=(l_{t-1}+0.85 b_{t-1}+s_{t-m})(1+\epsilon _t)$
$l_t=(l_{t-1}+0.85 b_{t-1}+0.612(l_{t-1}+0.85 b_{t-1}+s_{t-m})(\epsilon _t)$
$b_t=(0.85 b_{t-1}+0.155(l_{t-1}+0.85 b_{t-1}+s_{t-m})(\epsilon _t)$
$s_t=s_{t-m}+0.191l_{t-1}+0.85 b_{t-1}+s_{t-m})(\epsilon _t)$

```{r}
ets.model <- ets(solar_production_train, lambda = 'auto')
summary(ets.model)
checkresiduals(ets.model)

```

The ETS model that produces by far the lowest information criterion is the "AAA" model after Box-Cox transformation; unfortunately, we're still seeing significant autocorrelation in the residuals.

# Performing CV and calculating MAE, MSE, and AICc for ETS AAA:

```{r}
k <- 180 # Minimum data length for fitting a model
n <- length(solar_production_train) # Number of data points
p <- 12 # Period
H <- 12 # Forecast Horizon

st <- tsp(solar_production_train)[1]+(k-2)/p # start time in units
print(st)
model_error <- matrix(NA,n-k,H)
AICc_model <- list()

### One Month rolling forecasting ###
  
# Sliding Window - With window of length = 180

for(i in 1:(n-k))
{

  train_sli <- window(solar_production_train, start=st+(i-k+1)/p, end=st+i/p) ## Window Length: k
  val <- window(solar_production_train, start=st + (i+1)/p, end=st + (i+H)/p) ## Window Length: H
  if (i<4) {
  cat(c("*** CV", i,":","len(Sliding Window):",length(train_sli), "len(Val):",length(val),'\n'  ))
  cat(c("*** TRAIN - Sliding WIndow:",tsp(train_sli)[1],'-',tsp(train_sli)[2],'\n'))
  cat(c("*** TEST:",tsp(val)[1],'-',tsp(val)[2],'\n'))
  cat("*************************** \n \n")
  }
  
  fit_model <- ets(train_sli, lambda = 'auto')
  
  fcast_model <- forecast(fit_model, h=H)
  
  model_error[i,1:length(val)] <- fcast_model[['mean']]-val
  
  AICc_model <- append(AICc_model, fit_model$aicc)
}
```
```{r}
MAE <- function(x){
  return(mean(abs(x),na.rm=TRUE))
}

MSE <- function(x){
  return(mean(x^2, na.rm=TRUE))
}

flattenlist <- function(x){  
  morelists <- sapply(x, function(xprime) class(xprime)[1]=="list")
  out <- c(x[!morelists], unlist(x[morelists], recursive=FALSE))
  if(sum(morelists)){ 
    Recall(out)
  }else{
    return(out)
  }
}

MAE_model <- apply(model_error, 1, MAE)
MSE_model <- apply(model_error, 1, MSE)

ets.MAE <- mean(MAE_model)
ets.MSE <- mean(MSE_model)
results <- cbind(results,rbind(ets.MAE, ets.MSE))
colnames(results) <- c("Baseline","ARIMA(4,1,4)(0,1,1)12", "ARIMA(0,1,2)(0,1,1)12", "ets")

cat("Mean AICc:", mean(unlist(flattenlist(AICc_model))))
```
```{r}
results

```

# Forecasting Error for ETS:

```{r}
ets.forecasting.MSE <- mean((forecast(ets.model, h = 12)$mean - solar_production_test)^2)
ets.forecasting.MSE
ets.forecasting.MAE <- mean(abs(forecast(ets.model, h = 12)$mean - solar_production_test))
ets.forecasting.MAE
plot(forecast(ets.model, h = 12))

```

# TBATS:

```{r}
tbats.model <- tbats(solar_production_train, use.damped.trend = FALSE)
tbats.model
checkresiduals(tbats.model)
```

# Performing CV and calculating MAE, MSE, and AICc for TBATS:

```{r}
 
k <- 220 # Minimum data length for fitting a model
n <- length(solar_production_train) # Number of data points
p <- 12 # Period
H <- 12 # Forecast Horizon

st <- tsp(solar_production_train)[1]+(k-2)/p # start time in units
print(st)
model_error <- matrix(NA,n-k,H)
AICc_model <- list()

### One Month rolling forecasting ###
  
# Sliding Window - With window of length = 180

for(i in 1:(n-k))
{

  train_sli <- window(solar_production_train, start=st+(i-k+1)/p, end=st+i/p) ## Window Length: k
  val <- window(solar_production_train, start=st + (i+1)/p, end=st + (i+H)/p) ## Window Length: H
  if (i<4) {
  cat(c("*** CV", i,":","len(Sliding Window):",length(train_sli), "len(Val):",length(val),'\n'  ))
  cat(c("*** TRAIN - Sliding WIndow:",tsp(train_sli)[1],'-',tsp(train_sli)[2],'\n'))
  cat(c("*** TEST:",tsp(val)[1],'-',tsp(val)[2],'\n'))
  cat("*************************** \n \n")
  }
  
  fit_model <- tbats(train_sli, use.damped.trend = FALSE)
  
  fcast_model <- forecast(fit_model, h=H)
  
  model_error[i,1:length(val)] <- fcast_model[['mean']]-val
  
  AICc_model <- append(AICc_model, fit_model$aicc)
}
```
```{r}
MAE <- function(x){
  return(mean(abs(x),na.rm=TRUE))
}

MSE <- function(x){
  return(mean(x^2, na.rm=TRUE))
}

flattenlist <- function(x){  
  morelists <- sapply(x, function(xprime) class(xprime)[1]=="list")
  out <- c(x[!morelists], unlist(x[morelists], recursive=FALSE))
  if(sum(morelists)){ 
    Recall(out)
  }else{
    return(out)
  }
}

MAE_model <- apply(model_error, 1, MAE)
MSE_model <- apply(model_error, 1, MSE)

tbats.MAE <- mean(MAE_model)
tbats.MSE <- mean(MSE_model)
results <- cbind(results,rbind(tbats.MAE, tbats.MSE))
colnames(results) <- c("Baseline","ARIMA(4,1,4)(0,1,1)12", "ARIMA(0,1,2)(0,1,1)12", "ets", "tbats")

cat("Mean AICc:", mean(unlist(flattenlist(AICc_model))))
```
```{r}
results


```
results
# Forecasting Error for TBATS:

```{r}
tbats.forecasting.MSE <- mean((forecast(tbats.model, h = 12)$mean - solar_production_test)^2)
tbats.forecasting.MSE
tbats.forecasting.MAE <- mean(abs(forecast(tbats.model, h = 12)$mean - solar_production_test))
tbats.forecasting.MAE
plot(forecast(tbats.model, h = 12))
```

# Regression with ARIMA errors:
solar[1:348]
```{r}

train <- ts(solar[1:348,],start=c(1989,1), frequency=12)
test <- ts(solar[349:360,], end=c(2018,12), frequency=12)
head(train)
```

```{r}

lin_model <- glm(train[,2]~train[,3]+train[,9]+train[,10],data=train)
summary(lin_model)
tsdisplay(ts(lin_model$residuals))
```

```{r}
regressionArima = auto.arima(ts(lin_model$residuals, frequency = 12),lambda = "auto") 
FullArima <- Arima(train[,2],order = c(2,1,2), seasonal = c(1,1,1), lambda = "auto", xreg = exojeans)
summary(FullArima)
```

```{r}
RegX_forecast_smart <- forecast(FullArima,xreg = exojeans_smart, h = 12)
summary(RegX_forecast_smart)
```

```{r}
RegX_forecast_naive <- forecast(FullArima,xreg = exojeans_naive, h = 12)
summary(RegX_forecast_naive)
```

```{r}
(Smart_ARIMAX_MSE <- mean((RegX_forecast_smart$mean - test[,2])^2))
(Smart_ARIMAX_MAE <- mean(abs(RegX_forecast_smart$mean - test[,2])))
(Naive_ARIMAX_MSE <- mean((RegX_forecast_naive$mean - test[,2])^2))
(Naive_ARIMAX_MAE <- mean(abs(RegX_forecast_naive$mean - test[,2])))
```

```{r}
plot(RegX_forecast_naive, type = "l", col = 1, ylab = "Solar Production", xlab = "Year", main = "Regression with ARIMA Errors (Naive)")
plot(RegX_forecast_smart, type = "l", col = 1, ylab = "Solar Production", xlab = "Year", main = "Regression with ARIMA Errors (Smart)")
legend("topleft",legend=c("Naive Reg","'Smart' Reg", "True Value"),col=1:3,lty=1)
```

